From Reversible Programs to \\ Univalent Universes and Back

### 2021-10-27 Pi/paper/ufin.tex
Expand homotopy types and add fib/im
Fixes #149

A homotopy between functions $f, g : A \to B$, written $f \htpy g$, is given by pointwise equality between them,
$\dfun*{x:A}{f(x) \id_{B} g(x)}$. The identity type for functions is equivalent to homotopies between them ${(f \id_{A
        \to B} g)} \eqv {(f \htpy g)}$, by $\term{funext}$ and $\term{happly}$. An equivalence between types $A \eqv B$ is given
by a pair of functions between them which compose to the identity, $f \comp g \htpy \idfunc_{B}$ and $g \comp f \htpy
  \idfunc_{A}$ (also see~\cref{prop:three-chars-equiv} in the appendix), and this is equivalent to the identity type for
the universe, $(A \id_{\UU} B) \eqv (A \eqv B)$, which is the univalence principle.

@@

\begin{toappendix}
  \begin{definition}[${\fib_{f}} : B \to \UU$]
    \label{def:fib}
    The fiber of $f : A \to B$ at $b : B$ is
    \[
      \fib_{f}(b) \defeq \dsum*{a:A}{f(a) \id_{B} b}.
    \]
  \end{definition}

  \begin{definition}[${\term{im}} : (f : A \to B) \to \UU$]
    \label{def:im}
    The image of $f$ is the (-1)-truncation of its fiber.
    \[
      \im{f} \defeq \dsum*{b:B}{\Trunc[-1]{\fib_{f}(b)}}
    \]
  \end{definition}

  \begin{proposition}
    \label{prop:three-chars-equiv}
    The following are equivalent.
    \begin{enumerate}
      \item $f : A \to B$ is an equivalence.
      \item $f$ has a left and right inverse.
      \item $f$ has contractible fibers.
    \end{enumerate}
  \end{proposition}
\end{toappendix}

@@

>>>>

A homotopy between functions $f, g : A \to B$, written $f \htpy g$, is given by pointwise equality between them,
$\dfun*{x:A}{f(x) \id_{B} g(x)}$. The identity type for functions is equivalent to homotopies between them ${(f \id_{A
\to B} g)} \eqv {(f \htpy g)}$, by $\term{funext}$ and $\term{happly}$. An equivalence between types $A \eqv B$ is given
by a pair of functions between them which compose to the identity, $f \comp g \htpy \idfunc_{B}$ and $g \comp f \htpy
\idfunc_{A}$ (also see~\cref{prop:three-chars-equiv}), and this is equivalent to the identity type for the universe, $(A
\id_{\UU} B) \eqv (A \eqv B)$, which is the univalence principle.

@@

@@

\noindent
Given a function $f : A \to B$, there is an associated fibration over $B$, given by the homotopy fiber. The image of $f$
is defined to be the (-1)-truncation of its fiber.
\begin{gather*}
  \fib_{f}(b) \defeq \dsum*{a:A}{f(a) \id_{B} b}
  \qquad
  \im{f} \defeq \dsum*{b:B}{\Trunc[-1]{\fib_{f}(b)}}
\end{gather*}

\begin{proposition}
  \label{prop:three-chars-equiv}
  The following propositions are equivalent.
  \begin{enumerate}
    \item $f : A \to B$ is an equivalence.
    \item $f$ has a left and right inverse.
    \item $f$ has contractible fibers.
  \end{enumerate}
\end{proposition}
